Software Citation Principles

### 2016-07-21 sc-principles/software-citation-principles.tex
edits to improve structure and flow

The group gathered members (see Appendix~\ref{app:wg_members}) in April and May 2015,
and then began work in June, with a number of meetings
and some off-line work by group members to gather materials documenting existing
practices in member disciplines; gather materials from workshops and other reports;
review those materials, identifying overlaps and differences; and subsequently draft an initial version of this document.
The draft Software Citation Principles document was discussed in a day-long workshop and presented at the FORCE2016 Conference~\cite{force2016} in April 2016.
Members of the workshop and greater FORCE11 community gave feedback, which we recorded here in Appendix~\ref{app:force16-feedback}.
This discussion led to some changes in the use cases and discussion, although the principles themselves were not modified.
We also plan to initiate a follow-on working group that will work with stakeholders to ensure that these principles impact the research process.

The software citation principles were originally based on an adaptation of the
FORCE11 Data Citation Principles~\cite{data-citation-principles}, and
modified based on discussions of the FORCE11 Software Citation Working
Group (see Appendix~\ref{app:wg_members} for members), information from the use
cases in \S\ref{sec:use_cases}, and the related work in \S\ref{sec:related_work}.
The adaptations were made because software, while similar to data in terms of
not traditionally having been cited in publications, is also different than
data. Software can be used to express or explain concepts, it is typically
updated more frequently than data\footnote{Datasets can of course be frequently
updated, but each update likely represents the findings of a new or repeated
study.}, and it is executable.
\niemnote{Dan: can you add more here?}
Also, while software can be considered a type of data, the converse is not
generally true.

>>>>

The group gathered members (see Appendix~\ref{app:wg_members}) in April and
May 2015, and then began work in June. This materialize as a number of meetings
and offline work by group members to document existing
practices in member disciplines; gather materials from workshops and other
reports; review those materials, identifying overlaps and differences;
create a list of use cases related to software citation, recorded in
Appendix~\ref{app:usecases}; and subsequently draft an initial version of this document.
The draft Software Citation Principles document was discussed in a day-long
workshop and presented at the FORCE2016 Conference~\cite{force2016} in April 2016.
Members of the workshop and greater FORCE11 community gave feedback, which we
recorded here in Appendix~\ref{app:force16-feedback}. This discussion led to
some changes in the use cases and discussion, although the principles themselves
were not modified. We also plan to initiate a follow-on working group that will
work with stakeholders to ensure that these principles impact the research process.

The process of creating the software citation principles began by adapting
the FORCE11 Data Citation Principles~\cite{data-citation-principles}. These
were then modified based on discussions of the FORCE11 Software Citation
Working Group (see Appendix~\ref{app:wg_members} for members), information from
the use cases in \S\ref{sec:use_cases}, and the related work in
\S\ref{sec:related_work}.
The adaptations were made because software, while similar to data in terms of
not traditionally having been cited in publications, is also different than
data. Software can be used to express or explain concepts, it is typically
updated more frequently than data\footnote{Datasets can of course be frequently
updated, but each update likely represents the findings of a new or repeated
study.}, and it is executable.
\niemnote{Dan: can you add more here?}
Also, while software can be considered a type of data, the converse is not
generally true.
